---
title: "Horror Movies"
author: "Matthew"
date: "2022-11-01"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, dpi = 500)
library(tidyverse)
library(scales)
library(tidymodels)
library(lubridate)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
```

```{r}
horror_movies <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-11-01/horror_movies.csv')
```

# EDA

```{r}
horror_movies %>% 
  count(adult, sort = TRUE)
```

The adult variable only contains on distinct outcome (*FALSE*) and thus provides no information.

```{r}
horror_movies <- 
  horror_movies %>% 
    select(-adult)
```


## Vote Average

```{r}
horror_movies %>% 
  ggplot(aes(vote_average)) + geom_histogram() +
  labs(title = "Count of Vote Averages", y = "", x = "")
```

```{r}
horror_movies %>% 
  separate_rows(genre_names, sep = ", ") %>% 
  filter(genre_names != "Horror") %>% 
  group_by(genre_names) %>% 
  summarize(avg = mean(vote_average)) %>% 
  ggplot(aes(avg, fct_reorder(genre_names, avg))) + geom_col(fill = "skyblue", color = "black") +
  labs(y = "", x = "", title = "Vote Average by Sub-Genre") + geom_text(aes(label = round(avg, 3)), nudge_x = -0.3)
```

```{r}
horror_movies %>% 
  separate_rows(genre_names, sep = ", ") %>% 
  mutate(genre_names = fct_lump(genre_names, 10),
         genre_names = ifelse(genre_names == "Horror", "All Horrors", as.character(genre_names))) %>% 
  filter(genre_names != "Other") %>% 
  group_by(genre_names) %>% 
  summarize(avg = mean(vote_average)) %>% 
  ggplot(aes(avg, fct_reorder(genre_names, avg))) + geom_col(aes(fill = genre_names == "All Horrors"), color = "black") +
  labs(y = "", x = "", title = "Vote Average of 10 Most Common Sub-Genres") + geom_text(aes(label = round(avg, 3)), nudge_x = -0.3) +
  theme(legend.position = "") + scale_fill_manual(values = c("darkorange", "purple"))
```

### Films with 0 Vote Average

```{r}
horror_movies %>% 
  filter(vote_average == 0) %>% 
  count(vote_count, sort = TRUE)
```

Majority of films with a *vote_average* of 0 have very few votes, with the overwhelming majority being 0 votes.

## Budget

```{r}
horror_movies %>% 
  ggplot(aes(budget)) + geom_histogram() + scale_x_log10(labels = comma_format()) +
  labs(title = "Budget", y = "", x = "")
```

```{r}
horror_movies %>% 
  count(budget, sort = TRUE)
```

As is the case with vote count, *budget* also has many 0's.

```{r}
horror_movies %>% 
  filter(budget == 0 & vote_count == 0) %>% 
  summarize(n())
```

Over **10,000** rows of data contain both a *budget* of 0 and *vote_count* of 0. This leads me to believe a lot of the data is actually missing.

```{r}
horror_movies %>% 
  filter(budget == 0 | vote_count == 0) %>% 
  ggplot(aes(budget, vote_count)) + geom_point() +
  geom_text(aes(label = title), check_overlap = TRUE, size = 3, hjust = "inward", nudge_x = 2.5e6,
            data = horror_movies %>% filter((budget > 0 & vote_count == 0) | (budget == 0 & vote_count > 0))) + 
  labs(title = "Movies with Vote Count or Budget Equals 0") 
```

```{r}
horror_movies %>% 
  mutate(zero_counts = ifelse(vote_count == 0, "Yes", "No")) %>% 
  group_by(y = year(release_date)) %>% 
  count(zero_counts) %>% 
  ggplot(aes(y, n, color = zero_counts)) + geom_line() + scale_color_brewer(palette = "Set1", direction = -1) +
  labs(title = "Number of Movies with no Votes over time")
```

```{r}
horror_movies %>% 
  mutate(budget = ifelse(budget == 0, "Yes", "No")) %>% 
  group_by(y = year(release_date)) %>% 
  count(budget) %>% 
  ggplot(aes(y, n, color = budget)) + geom_line() + scale_color_brewer(palette = "Set1", direction = -1) +
  labs(title = "Number of Movies with no Budget over time")
```

```{r}
horror_movies %>% 
  filter(vote_count != 0, budget != 0) %>% 
  ggplot(aes(budget, vote_count)) + geom_point(alpha = 0.5, color = "purple") + geom_smooth(se = FALSE) +
  scale_x_log10(labels = comma_format()) + scale_y_log10(labels = comma_format(accuracy = 1)) + 
  labs(y = "", x = "", title = "Vote Count by Budget")
```

# Halloween

```{r}
horror_movies %>% 
  mutate(month = month(release_date),
         day = day(release_date)) %>% 
  ggplot(aes(month)) + geom_bar(aes(fill = (month == 10))) + 
  scale_x_continuous(breaks = seq(1,12)) + scale_fill_manual(values = c("grey20","darkorange3")) +
  theme(panel.grid.minor = element_blank(), legend.position = "") + 
  labs(y = "", x = "Month", title = "Horror Movies Released in October vs Other Months")

horror_movies %>% 
  mutate(month = month(release_date), day = day(release_date)) %>% 
  summarize(Halloween = mean(month == 10 & day == 31)*100, October = mean(month == 10)*100)
```

## October

```{r}
horror_movies %>% 
  filter(month(release_date) == 10) %>% 
  count(day(release_date)) %>% 
  ggplot(aes(`day(release_date)`, n)) + geom_col(color = "black", fill = "darkorange") +
  geom_label(aes(x = 26, y = 500, label = "Halloween"), fill = "darkorange", fontface = "bold") +
  scale_x_continuous(breaks = seq(1, 31, 3)) +
  geom_curve(aes(x = 26, y = 460, xend = 30, yend = 450), curvature = 0.3, color = "darkred", size = 1,
             arrow = arrow(length = unit(0.03, "npc"))) + 
  theme(panel.grid = element_blank(), plot.title = element_text(hjust = 0.5), 
        panel.background = element_rect(fill = "black", color  =  NA)) +
  labs(title = "Horror Movie Release Day in October", y = "", x = "") 
```

