---
title: "Boardgames"
author: "Matthew"
date: "1/25/2022"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(ggthemes)
library(tidytext)
theme_set(theme_bw())
ratings <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-25/ratings.csv')
details <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-25/details.csv')

games <- ratings %>%
  inner_join(details, by = "id") %>%
  rename_with(~gsub(pattern = "boardgame",replacement = "", .x))

games <- games %>%
  select(-primary, -thumbnail, -url)
```


# EDA
## Check Data
```{r}
games %>%
  select_if(is.numeric) %>%
  gather() %>%
  ggplot(aes(value)) + geom_histogram() + facet_wrap(~key, scales = "free")

games %>%
  count(mechanic, sort = TRUE)
```


## Facet numerics
### Playtime
```{r}
games %>%
  group_by(name) %>%
  arrange(-average) %>%
  head(20) %>%
  pivot_longer(contains("playtime"), names_to = "key", values_to = "value") %>%
  ggplot(aes(average, fct_reorder(name, average), fill = factor(value))) + geom_col(color = "black") + facet_wrap(~key) +
  scale_fill_viridis_d(option = "magma") +
  labs(x = "Average Rating", y = "", title = "Top 20 Games") 
```


### Number of players
```{r}
games %>%
  group_by(name) %>%
  arrange(-average) %>%
  head(20) %>%
  pivot_longer(contains("player"), names_to = "key", values_to = "value") %>%
  ggplot(aes(average, fct_reorder(name, average), fill = factor(value))) + geom_col(color = "black") + facet_wrap(~key) +
  scale_fill_viridis_d(option = "magma") +
  labs(x = "Average Rating", y = "", title = "Top 20 Games")
```


### Gplot column function for filling by numeric
```{r}
Gplot <- function(x) {
  games %>%
    group_by(name) %>%
    arrange(-average) %>%
    head(20) %>%
    ggplot(aes(average, fct_reorder(name, average), fill = {{x}})) + geom_col(color = "black") +
    labs(x = "Average Rating", y = "", title = "Top 20 Games") 
}

Gplot(maxplayers)
Gplot(minplayers)
```


### Gplot2 point function of numerics and average
```{r}
Gplot2 <- function(x) {
  games %>%
    filter({{x}} != 0) %>%
    ggplot(aes({{x}}, average)) + geom_point(alpha = 0.1) + scale_x_log10() + 
    geom_hline(yintercept = mean(games$average), lty = 3, size = 1.5, color = "red") +
    geom_smooth()
}

Gplot2(wanting)
Gplot2(wishing)
Gplot2(trading)
Gplot2(owned)
```


## Game Mechanics
```{r}
games %>%
  unnest_tokens(word, mechanic) %>%
  group_by(word) %>%
  summarize(n = n(), mean = mean(average)) %>%
  arrange(-n) %>%
  filter(!word %in% c("game", "and", "player"), !is.na(word)) %>%
  head(30) %>%
  ggplot(aes(mean, fct_reorder(word, mean), fill = word)) + geom_col() + geom_vline(xintercept = mean(games$average), lty = 3) +
  theme(legend.position = "") + labs(y = "", x = "Average rating", title = "Average rating by most common game mechanics")
```

### Tidy mechanic column and furthur analysis
```{r}
games %>%
  separate_rows(mechanic, sep = ",") %>%
  mutate(mechanic = str_replace_all(mechanic, "\\[|\\]", ""),
         mechanic = str_replace_all(mechanic, "\'", ""),
         mechanic = str_trim(mechanic, "both")) %>% 
  group_by(mechanic) %>%
  summarize(n = n(), average = mean(average)) %>%
  filter(n >100) %>%
  arrange(-average) %>%
  head(30) %>%
  ggplot(aes(average, fct_reorder(mechanic, average), fill = average)) + geom_col(color = "black") + 
  theme(legend.position = "") +
  geom_vline(xintercept = mean(games$average), lty = 3, color = "red") +
  labs(y = "", x = "", title = "Average rating by common mechanics", subtitle = "Highest 30 Mechanics",
       caption = "Red line indicates global average")


games %>%
  separate_rows(mechanic, sep = ",") %>%
  mutate(mechanic = str_replace_all(mechanic, "\\[|\\]", ""),
         mechanic = str_replace_all(mechanic, "\'", ""),
         mechanic = str_trim(mechanic, "both")) %>% 
  group_by(mechanic) %>%
  summarize(n = n(), average = mean(average)) %>%
  filter(n >100, !is.na(mechanic)) %>%
  arrange(average) %>%
  head(30) %>%
  ggplot(aes(average, fct_reorder(mechanic, average), fill = average)) + geom_col(color = "black") + 
  geom_vline(xintercept = mean(games$average), lty = 3, color = "red") +
  theme(legend.position = "") +
  labs(y = "", x = "", title = "Average rating by common mechanics", subtitle = "Lowest 30 Mechanics",
       caption = "Red line indicates global average")



games %>%
  separate_rows(mechanic, sep = ",") %>%
  mutate(mechanic = str_replace_all(mechanic, "\\[|\\]", ""),
         mechanic = str_replace_all(mechanic, "\'", ""),
         mechanic = str_trim(mechanic, "both")) %>% 
  group_by(mechanic) %>%
  summarize(n = n(), average = mean(average)) %>%
  filter(n >100, !is.na(mechanic)) %>%
  arrange(-average) %>%
  ggplot(aes(average, n)) + geom_text(aes(label = mechanic), check_overlap = TRUE)
```


## Game category
```{r}
games %>%
  separate_rows(category, sep = ",") %>%
  mutate(category = str_replace_all(category, "\\[|\\]", ""),
         category = str_replace_all(category, "\'",""),
         category = str_replace_all(category, "\\\"",""),
         category = str_trim(category, "both")) %>%
  group_by(category) %>%
  drop_na() %>%
  summarize(n = n(), mean = mean(average)) %>%
  arrange(-mean) %>%
  filter(n > 25) %>%
  head(25) %>%
  ggplot(aes(mean, fct_reorder(category, mean), fill = mean)) + geom_col(color = "black") + 
  geom_vline(xintercept = mean(games$average), lty = 3, color = "red") +
  labs(y = "", x = "Average Rating", title = "Top 25 Categories", caption = "Only categories with more than 25 occurances") + 
  theme(legend.position = "")

games %>%
  separate_rows(category, sep = ",") %>%
  mutate(category = str_replace_all(category, "\\[|\\]", ""),
         category = str_replace_all(category, "\'",""), 
         category = str_replace_all(category, "\\\"",""),
         category = str_trim(category, "both")) %>%
  group_by(category) %>%
  drop_na() %>%
  summarize(n = n(), mean = mean(average)) %>%
  arrange(mean) %>%
  filter(n > 10) %>%
  head(25) %>%
  ggplot(aes(mean, fct_reorder(category, mean), fill = mean)) + geom_col(color = "black") + 
  geom_vline(xintercept = mean(games$average), lty = 3, color = "red") +
  labs(y = "", x = "Average Rating", title = "Bottom 25 Categories", caption = "Only categories with more than 10 occurances") +
  theme(legend.position = "")
```


## Bayes Average and gplot function
```{r}
gplot <- function(x,y) {
  games %>%
    filter({{x}} > 0) %>%
    group_by({{x}}) %>%
    summarize(y = mean({{y}})) %>%
    ggplot(aes({{x}}, y)) + geom_point()
}
gplot(wanting, bayes_average) + scale_x_sqrt()
```


### Gathered Bayes/Average and make another gplot function
```{r}
gplot2 <- function(x) {
  games %>% 
    gather(key = key, value = value, bayes_average, average) %>%
    filter({{x}} > 0) %>%
    group_by({{x}}, key) %>%
    summarize(value = mean(value)) %>%
    ggplot(aes({{x}}, value)) + geom_point(alpha = 0.3) + facet_wrap(~key)
}

gplot2(wishing) + scale_x_log10() + geom_smooth(method = "lm") + 
  labs(title = "Comparison of Average vs Bayes Average", subtitle = "X is on a log10 scale")
gplot2(wishing) + scale_x_sqrt() + geom_smooth(method = "lm") + 
  labs(title = "Comparison of Average vs Bayes Average", subtitle = "X is on a square root scale")
```


#### Quick lm models
```{r}
lm(bayes_average ~ sqrt(wishing) * sqrt(wanting) + log(minplayers), games %>% filter(wishing >0, minplayers >0)) %>% summary()
lm(average ~ log(wishing), games %>% filter(wishing >0)) %>% summary()
```


##### Plots of tidy lm models
```{r}
lm(bayes_average ~ sqrt(wishing) * sqrt(wanting), games %>% filter(wishing >0)) %>% 
  tidy(conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, term, color = term)) + geom_point() + geom_errorbar(aes(xmin = conf.low, xmax = conf.high))

lm(bayes_average ~ ., games %>% select_if(is.numeric) %>% 
  filter(wishing >0)) %>% 
  tidy(conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>% 
  drop_na() %>%
  ggplot(aes(estimate, fct_reorder(term, estimate), color = term)) + 
  geom_point() + geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
  theme(legend.position = "") + labs(y = "", title = "Linear model using only numeric estimates")


```


```{r}
library(tidymodels)
library(mgcv)
games_split <- initial_split(games, strata = bayes_average)
games_train <- training(games_split)
games_test <- testing(games_split)

set.seed(2022)

model <- lm(bayes_average ~ sqrt(wishing) * sqrt(wanting) + minplayers + average, games_train)

model %>% summary()

games_test %>%
  mutate(predictions = predict(model, games_test),
         residuals = bayes_average - predictions) %>%
  ggplot(aes(predictions, residuals)) + geom_point(alpha = 0.1) + 
  geom_hline(yintercept = 0, lty = 2, color = "blue", size = 1) +
  labs(title = "Residual plot for first lm")

games_test %>%
  mutate(predictions = predict(model, games_test)) %>%
  group_by(wishing) %>%
  summarize(bayes_average = mean(bayes_average),
            predictions = mean(predictions)) %>%
  ggplot(aes(wishing)) + geom_point(aes(y = bayes_average), alpha = 0.3) + 
  geom_point(aes(y = predictions), color = "blue", alpha = 0.3) + scale_x_sqrt()

games_test %>%
  mutate(predictions = predict(model, games_test)) %>%
  ggplot(aes(predictions, bayes_average)) + geom_point(alpha = 0.07) + 
  geom_abline(color = "blue", linetype = "dashed", size = 1) +
  xlim(5.5,8.5) + ylim(5.5,8.5)
  
```


##### Gam model
```{r}
model2 <- gam(bayes_average ~ s(wishing), family = gaussian, data = games_train)

model2 %>% summary()

games_test %>%
  mutate(predictions = predict(model2, games_test),
         residuals = bayes_average - predictions) %>%
  ggplot(aes(predictions, residuals)) + geom_point(alpha = 0.1) + 
  geom_hline(yintercept = 0, lty = 2, color = "blue", size = 1) +
  labs(title = "Residual plot for first Gam")

games_test %>%
  mutate(predictions = predict(model2, games_test)) %>%
  group_by(wishing) %>%
  summarize(bayes_average = mean(bayes_average),
            predictions = mean(predictions)) %>%
  ggplot(aes(wishing)) + geom_point(aes(y = bayes_average), alpha = 0.3) + 
  geom_point(aes(y = predictions), color = "blue", alpha = 0.3) + scale_x_sqrt()

games_test %>%
  mutate(predictions = predict(model2, games_test)) %>%
  ggplot(aes(predictions, bayes_average)) + geom_point(alpha = 0.07) + 
  geom_abline(color = "blue", linetype = "dashed", size = 1) +
  xlim(5.5,8.5) + ylim(5.5,8.5)
```

##### Both models
```{r}
games_test %>%
  mutate(predictions = predict(model2, games_test),
         model = "2") %>%
  bind_rows(games_test %>%
              mutate(predictions = predict(model, games_test),
                     model = "1")) %>%
  group_by(wishing, model) %>%
  summarize(bayes_average = mean(bayes_average),
            predictions = mean(predictions)) %>%
  ggplot(aes(wishing)) + geom_point(aes(y = bayes_average), alpha = 0.3) + 
  geom_point(aes(y = predictions, color = model), alpha = 0.3) + scale_x_sqrt() +
  labs(title = "Both models fitted", subtitle = "Square root scale")

```


```{r}
games_df <- games %>%
  select(bayes_average, category, wishing) %>%
  drop_na() %>%
  separate_rows(category, sep = ",") %>%
  mutate(category = str_replace_all(category, "\\[|\\]", ""),
         category = str_replace_all(category, "\'",""), 
         category = str_replace_all(category, "\\\"",""),
         category = str_trim(category, "both")) %>%
  mutate_if(is.character, factor) %>%
  mutate(category = fct_lump(category, prop = 0.01))
```


#### Gam model 3 using categories
```{r}
df_split <- initial_split(games_df, strata = bayes_average)
df_train <- training(df_split)
df_test <- testing(df_split)

model3 <- gam(bayes_average ~ s(wishing) + category, gaussian, df_train)

model3 %>% summary()

df_fit <- df_test %>%
  mutate(predictions = predict(model3, df_test),
         residuals = bayes_average - predictions)

df_fit %>%
  group_by(wishing) %>%
  summarize(bayes_average = mean(bayes_average),
            predictions = mean(predictions)) %>%
  ggplot(aes(wishing)) + geom_point(aes(y = bayes_average), alpha = 0.3) + 
  geom_point(aes(y = predictions), color = "blue", alpha = 0.3) + scale_x_sqrt() +
  labs(title = "GAM model", subtitle = "Includes wishing and top 31 most common categories")
```

#### Model 2
```{r}
games_test %>%
  filter(bayes_average > 0) %>%
  mutate(predictions = predict(model2, games_test %>% filter(bayes_average > 0)),
         residuals = bayes_average - predictions) %>%
  summarize(rmse = sqrt(mean(residuals^2)),
            relerr = sqrt(mean((residuals/bayes_average)^2)))
```

#### Model 3
```{r}
df_test %>%
  filter(bayes_average > 0) %>%
  mutate(predictions = predict(model3, df_test %>% filter(bayes_average > 0)),
         residuals = bayes_average - predictions) %>%
  summarize(rmse = sqrt(mean(residuals^2)),
            relerr = sqrt(mean((residuals/bayes_average)^2)))
```



