---
title: "Weather Forcasts"
author: "Matthew"
date: "2022-12-20"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, dpi = 500)
library(tidyverse)
library(scales)
library(patchwork)
library(tidymodels)
library(lubridate)
library(astsa)
library(forecast)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
```

```{r include=FALSE}
weather_forecasts <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-20/weather_forecasts.csv')
cities <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-20/cities.csv')
outlook_meanings <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-20/outlook_meanings.csv')
```

# Data Cleaning

```{r}
weather_forecasts <- weather_forecasts %>% 
  left_join(outlook_meanings) %>% 
  left_join(cities)
```

The *weather forecasts* dataset contains abbreviated weather outlooks, while the *outlook meanings* dataset contains non-abbreviated weather outlooks. The *weather forecasts* dataset and the *outlook meanings* dataset were joined together using the "forecast_outlook" column as a common key. This resulted in a combined dataset that includes both the abbreviated and non-abbreviated weather outlooks from the two datasets.

```{r}
weather_forecasts %>% 
  count(meaning, sort = TRUE)
```

## Data View

```{r}
weather_forecasts %>% 
  head(n = 10) %>% 
  knitr::kable()
```

## Missing Values

```{r}
skimr::skim(weather_forecasts) %>% select(skim_type, skim_variable, n_missing)
```

```{r}
weather_forecasts %>% 
  group_by(NA_meaning = is.na(meaning)) %>% 
  summarize(sum(is.na(forecast_temp)))

weather_forecasts %>% 
  group_by(forecast_hours_before) %>% 
  summarize(sum(is.na(forecast_temp)))
```

```{r}
weather_forecasts %>% 
  group_by(forecast_hours_before) %>% 
  summarize(sum(is.na(forecast_temp)))
```

Missing values for *forecast_temp* tend to appear where forecast *meaning* is also NA. They also slightly increase as the number of hours before. It is assumed that these missing values are due to there being no forecast at that time. Thus, these rows will not be dropped.

## Outliers

```{r}
weather_forecasts %>% 
  count(possible_error, sort = TRUE)
```

```{r}
weather_forecasts %>% 
  filter(possible_error == "observed_temp") %>% 
  select(contains("temp"))
weather_forecasts %>% 
  filter(possible_error == "forecast_temp") %>% 
  select(contains("temp"))
```

Column *possible_error* appears to show what **potentially** is flawed in that row, instead of inconsistancies between the forecast and observed. For instance, when *possible_error* equals "forecast_temp", all values for *forecast_temp* are **-10** and when it equals "observed_temp", values are either **0** or **108**.

# Exploratory Data Analysis

## Temperature Difference

```{r}
weather_forecasts <- weather_forecasts %>% 
  mutate(diff_temp = observed_temp - forecast_temp,
         prop_difftemp = diff_temp/observed_temp) 
```

Column of *temperature differences* and *proportional temperature differences* created above.

### States

```{r}
weather_forecasts %>% 
  filter(possible_error == "none") %>% 
  group_by(state) %>% 
  summarize(avg_diff = mean(abs(diff_temp), na.rm = TRUE)) %>% 
  slice_max(avg_diff, n = 10) %>% 
  ggplot(aes(avg_diff, fct_reorder(state, avg_diff), fill = avg_diff)) +
  geom_col(color = "black") + 
  scale_fill_viridis_c() +
  labs(x = "Average Forecast Temperature Error", y = "", 
       title = "Top 10 States with Highest Average Prediction Error") +
  theme(legend.position = "none")

```

```{r}
# Select top 6 states without highest average error
top6states <- weather_forecasts %>% 
  filter(possible_error == "none") %>% 
  group_by(state) %>% 
  summarize(avg_diff = mean(abs(diff_temp), na.rm = TRUE)) %>% 
  slice_max(abs(avg_diff), n = 6) %>% 
  pull(state)
```

```{r}
(weather_forecasts %>% 
  filter(state %in% top6states & possible_error == "none") %>% 
  ggplot(aes(date, diff_temp, color = state)) +
  geom_line() + theme(legend.position = "none") +
  labs(title = "Temperature Difference", x = "", y = "") +
  facet_wrap(~state, scales = "free")) /
(weather_forecasts %>% 
  filter(state %in% top6states & possible_error == "none") %>% 
  ggplot(aes(date, prop_difftemp, color = state)) + 
  geom_line() + theme(legend.position = "none") +
  labs(title = "Proportional Temperature Difference", x = "", y = "") +
  facet_wrap(~state, scales = "free"))
```

#### New York

```{r}
ny_weather <- weather_forecasts %>% 
  filter(state == "NY")
```

```{r}
ny_weather %>% 
  ggplot(aes(date, prop_difftemp)) +
  geom_line()
```

```{r}
ny_weather %>% 
  filter(prop_difftemp != -Inf) %>% 
  filter(abs(prop_difftemp) > 2.5) %>% 
  arrange(-abs(prop_difftemp)) %>% select(prop_difftemp)
```

# Time Series

Only keeping data within New York City, and since there are multiple values for each day, only the average will be kept.

```{r}
nycseries <- ny_weather %>% 
  filter(city == "NEW_YORK_CITY") %>% 
  distinct(date,observed_temp) %>% 
  group_by(date) %>% 
  summarize(observed_temp = mean(observed_temp, na.rm = TRUE)) %>% 
  arrange(date) %>% 
  drop_na()
```

## NYC Temperature

```{r}
(nycseries %>% 
  ggplot(aes(date, observed_temp)) +
  geom_line() +
  labs(y = "", x = "", title = "Observed Temperature")) /
(nycseries %>% 
  mutate(diff_temp = c(mean(nycseries$observed_temp, na.rm = TRUE), 
                       diff(observed_temp))) %>% 
  slice(-1) %>% 
  ggplot(aes(date, diff_temp)) +
  geom_line() +
  labs(y = "", x = "", title = "Differenced Temperature"))
```

### ARIMA

Using the *astsa* and *forcast* packages to fit an ARIMA model.

```{r}
# convert data to a timeseries object
nyctz <- ts(nycseries$observed_temp, start = c(2021,01,30), frequency = 365)
```

```{r}
auto.arima(nyctz)
sarima(nyctz, 2,1,3)
```

The *auto.arima* function suggests that the best ARIMA model is an ARIMA(2,1,3). Yet, it appears as though an ARIMA(1,1,3) is a better fit according to the diagnostics. Furthermore, the second AR term is not statistically significant. This model does not take into account any seasonality, although it is fairly clear that there is some seasonality in the temperature over time.

```{r}
sarima(nyctz, 1,1,3)
sarima.for(nyctz, 10, 1,1,3)
```

### GAM

```{r}
nycseries <- nycseries %>% 
  mutate(day = yday(date))

# future dates for extrapolation
fdates <- seq(as.Date("2022-05-30"), as.Date("2022-08-30"), by = 1) # future dates
```

```{r}
gam_mod <- mgcv::gam(observed_temp ~ s(day), 
                     data = nycseries, 
                     method = "REML")
cyclic_gam_mod <- mgcv::gam(observed_temp ~ s(day, bs = 'cp'), 
                     data = nycseries, 
                     method = "REML")

summary(gam_mod)
summary(cyclic_gam_mod)
```

For the GAM, the smooth term is set to a cyclic p-spline. This is done in order to connect the last day of the year to the first day of the year. When a non-cyclic spline is fit to the data, Jan 1st has a sharp vertical slope.

```{r}
(nycseries %>% 
  mutate(pred = cyclic_gam_mod$fitted.values) %>% 
  bind_rows(data.frame(date = fdates)) %>% 
  mutate(day = ifelse(is.na(day), yday(date), day),
         pred = ifelse(is.na(pred), predict(cyclic_gam_mod, cur_data()), pred)) %>% 
  ggplot(aes(date)) +
  geom_point(aes(y = observed_temp), color = "blue", alpha = 0.7) +
  geom_line(aes(y = pred), color = "red", size = 1) +
  geom_vline(xintercept = as.Date("2022-05-30"), linetype = "dashed") +
  labs(y = "", x = "", 
       title = "Cyclic P-Spline GAM")) /
(nycseries %>% 
  mutate(pred = gam_mod$fitted.values) %>% 
  bind_rows(data.frame(date = fdates)) %>% 
  mutate(day = ifelse(is.na(day), yday(date), day),
         pred = ifelse(is.na(pred), predict(gam_mod, cur_data()), pred)) %>% 
  ggplot(aes(date)) +
  geom_point(aes(y = observed_temp), color = "blue", alpha = 0.7) +
  geom_line(aes(y = pred), color = "red", size = 1) +
  geom_vline(xintercept = as.Date("2022-05-30"), linetype = "dashed") +
  labs(y = "", x = "", 
       title = "Basic Thin Plate GAM")) +
plot_annotation(title = "Temperature Modeled by Different Basis Functions")
```

### Polynomial Model

```{r}
lm_mod <- lm(observed_temp ~ I(day^4) + I(day^3) + I(day^2) + day, 
             data = nycseries)

summary(lm_mod)
```

Given the nature of this oscillating curve, it is quite easy to fit a polynomial model to fit the data. From the looks of it, this function appears to have 3 - 4 roots, and thus a 4 degree polynomial is fit.

```{r}
nycseries %>% 
  mutate(pred = predict(lm_mod, .)) %>% 
  bind_rows(data.frame(date = fdates)) %>% 
  mutate(day = ifelse(is.na(day), yday(date), day),
         pred = ifelse(is.na(pred), predict(lm_mod, cur_data()), pred)) %>% 
  ggplot(aes(date)) +
  geom_point(aes(y = observed_temp), color = "blue") +
  geom_line(aes(y = pred), color = "red") + 
  geom_vline(xintercept = as.Date("2022-05-30"), linetype = "dashed") +
  labs(y = "", x = "", title = "Temperature Modeled by Linear Regression")
```

It can be observed that the 4 degree polynomial is almost indistinguishable from the GAM fit. The only difference is that the GAM does not have the sharp decrease on Jan 1st because of its cyclic spline.
