---
title: "Weather Forcasts"
author: "Matthew"
date: "2022-12-20"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, dpi = 500)
library(tidyverse)
library(scales)
library(tidymodels)
library(lubridate)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
```

```{r include=FALSE}
weather_forecasts <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-20/weather_forecasts.csv')
cities <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-20/cities.csv')
outlook_meanings <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-20/outlook_meanings.csv')
```

# Data Cleaning

```{r}
weather_forecasts <- weather_forecasts %>% 
  left_join(outlook_meanings)  
```

The *weather forecasts* dataset contains abbreviated weather outlooks, while the *outlook meanings* dataset contains non-abbreviated weather outlooks. The *weather forecasts* dataset and the *outlook meanings* dataset were joined together using the "forecast_outlook" column as a common key. This resulted in a combined dataset that includes both the abbreviated and non-abbreviated weather outlooks from the two datasets.

```{r}
weather_forecasts %>% 
  count(meaning, sort = TRUE)
```

## Data View

```{r}
weather_forecasts %>% 
  head() %>% 
  knitr::kable()
```
